A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-12 docs/icccn2009/paper.tex
Abstract/Introduction
git-svn-id: svn://localhost/netbeams/trunk@370 559edde2-a4d4-44e0-96fc-4035e40d7396

\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

@@

Mention the nature of our sensor network: highly specialized sensors,
sparse topology, remote locations, long range communication.
\cite{roemer:2004}

Environmental monitoring depends on the reliable collection of
measurements from remotely deployed sensors and the rapid transfer of
those measurements to data centers where processing and analysis may
take place. Typically, sensors are deployed in out-of-the-way
locations, where physical access is limited and connections for
telemetry are poor or nonexistent. Maintaining the data stream from
these sensors places demands on human resources, requiring site visits
to service the sensors and to download data stored on internal memory.
Consequently, there is a lag of days to months between the time the
sensors perform the measurements and the time the measurements become
available. Advances in battery life and anti-fouling technology have
extended service cycles for field sensors, with the unwanted result of
further delaying access to monitoring data.  In addition, the
dependence on field site visits to alter of sensor characteristics, 
such as sampling rate, prevents adjustment in monitoring strategy in
response to rapidly developing events (e.g. oil spills, harmful algal
blooms).

To address these problems of data collection from remote
sensors, we propose to develop an automated end-to-end system based on
proven technology. It provides means to program and interrogate
sensors at the remote field locations and to transmit collected
measurements over long distances to data collections centers, where the
data is rapidly processed, archived, and made available to potential
users in near real-time.

@@

\section{Data Sensor Platform}

@@

\section{NetBEAMS}

@@

\begin{figure}[!htb]
\centering
\epsfig{file=dsp-messages-container, width=8cm}
\caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}
\end{figure}

@@

\section{Conclusions and Outlook}

>>>>

\maketitle
\begin{abstract}
  In this paper we describe a sensor network for environmental
  monitoring based on highly specialized, off-the-shelf sensing
  devices.  Environmental monitoring depends on the reliable
  collection of measurements from remotely deployed sensors and the
  rapid transfer of those measurements to data centers. Typically,
  sensors are deployed in out-of-the-way locations, where physical
  access is limited and connections for telemetry are poor or
  nonexistent. Maintaining the data stream from these sensors places
  demands on human resources, requiring site visits to service the
  sensors and to download data stored on internal memory.  The work
  described in this paper is a collaborative effort between the
  Computer Science Department at San Francisco State University and
  the Romberg Tiburon Center. The result of this collaboration is an
  environmental sensor network deployed in the San Francisco Bay. It
  provides means to program and interrogate sensors at the remote
  field locations and to transmit collected measurements to a remote
  database.
\end{abstract}

@@

A wireless sensor network consists of spatially distributed autonomous
devices using sensors to cooperatively monitor physical or
environmental conditions at different locations \cite{roemer:2004}.
Much research effort is focused on building wireless ad-hoc networks
where each sensor node participates in a multi-hop routing algorithm.
One the earliest examples of a wireless sensor network is the
SmartDust project \cite{smartdust:2001} where one objective was to
create autonomous sensing and communication in a cubic
millimeter-sized device.  While most wireless sensor networks focus on
small sensing devices such as the SmartDust mote, there exist a wide
variety of off-the-shelf commercially available sensors that are used
by environmental researchers. In contrast to the small sensing
devices, sensors for environmental monitoring are typically on the
order of one to two feet in length weighing several pounds. Our
focus is on building an end-to-end infrastructure for environmental
monitoring using specialized sensing devices.

Environmental monitoring depends on the reliable collection of
measurements from remotely deployed sensors and the rapid transfer of
those measurements to data centers. Typically, sensors are deployed in
out-of-the-way locations, where physical access is limited and
connections for telemetry are poor or nonexistent. Maintaining the
data stream from these sensors places demands on human resources,
requiring site visits to service the sensors and to download data
stored on internal memory.  Consequently, there is a lag of days to
months between the time the sensors perform the measurements and the
time the measurements become available. Advances in battery life and
anti-fouling technology have extended service cycles for field
sensors, with the unwanted result of further delaying access to
monitoring data.  In addition, the dependence on field site visits to
alter of sensor characteristics, such as sampling rate, prevents
adjustment in monitoring strategy in response to rapidly developing
events (e.g. oil spills, harmful algal blooms).

The work described in this paper is a collaborative effort between the
Computer Science Department at San Francisco State University and the
Romberg Tiburon Center (RTC), a research institute focusing on the
understanding of complex marine and estuarine environments. The result
of this collaboration is NetBEAMS (Networked Bay Environmental
Assessment Monitoring System), an environmental sensor network
deployed in the San Francisco Bay. NetBEAMS provides means to program
and interrogate sensors at the remote field locations and to transmit
collected measurements to a database, where the data is rapidly
processed, archived, and made available to potential users in near
real-time. In the following we give an overview of the NetBEAMS
architecture and implementation.  Section \ref{SEC_BACKGROUND}
describes a typical environmental sensor used by the RTC. Section
\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
to efficiently build up an end-to-end environmental sensor network.
Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
of the DSP. In Section \ref{SEC_CONCLUSION} we provide a conclusion
and outlook.

@@

\section{Data Sensor Platform}
\label{SEC_DSP}

@@

\section{NetBEAMS}
\label{SEC_NETBEAMS}

@@

\begin{figure}[!htb]
\centering
%\epsfig{file=dsp-messages-container, width=8cm}
\caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}
\end{figure}

@@

\section{Conclusions and Outlook}
\label{SEC_CONCLUSION}
